---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r read_data,results='hide'}
data1 <- read.csv("activity.csv")
data <- data1
```


## What is mean total number of steps taken per day?
```{r calculate mean, results='asis'}
steps_by_day <- tapply(data1$steps, data1$date, sum, na.rm=TRUE)
library(ggplot2)

a <- qplot(steps_by_day)
print(a)
mean_steps_by_day <- tapply(data1$steps, data1$date, mean, na.rm = TRUE)
median_steps_by_day <- tapply(data1$steps, data1$date, median, na.rm=TRUE)

library(xtable)
xtmean <- xtable(as.matrix(mean_steps_by_day))
print(xtmean, type="html")
xtmedian <- xtable(as.matrix(median_steps_by_day))
print(xtmedian, type = "html")


```


## What is the average daily activity pattern?

```{r averages}
library(ggplot2)
averages <- aggregate(x=list(steps = data$steps), by=list(interval=data$interval), FUN=mean, na.rm=TRUE)
ggplot(averages, aes(x=interval, y=steps)) +
    geom_line() +
    xlab("5-minute interval") + 
    ylab("average number of steps taken")


```

On average across all the days in the dataset, the 5-minute interval contains the maximum number of steps?
```{r, results='hide'}
averages[which.max(averages$steps),]

```

## Imputing missing values
-  Calculate and report the total number of missing values in the dataset
(i.e. the total number of rows with NAs)

```{r}
a <- data1[is.na(data1$steps),]
nrow(a)
```

Now we fill the missing values using the means of the 5-min intervals 
```{r filling data}
fill.val <- function(steps,interval) {
    filled <- NA
    if(!is.na(steps))
        filled <- c(steps)
    else 
        filled <- averages[averages$interval==interval, "steps"]
        return(filled)
}

filled.data <- data
filled.data$steps <- mapply(fill.val, filled.data$steps, filled.data$interval)

```
Now using the filled data set, lets make a histogram of the total number of steps taken each day and calculate the mean and the median total number of steps
```{r}
total.steps <- tapply(filled.data$steps, filled.data$date, FUN=sum)
qplot(total.steps, binwidth=1000, xlab="total number of steps taken each day")
mean(total.steps)
median(total.steps)
```
## Are there differences in activity patterns between weekdays and weekends?

```{r weekdays}
weekday.or.weekend <-function(date) {
    day <- weekdays(date)
    if(day %in% c("Monday","Tuesday", "Wednesday", "Thursday", "Friday"))
        return("weekday") 
    else if (day %in% c("Saturday", "Sunday"))
        return("weekend")
    else
        stop("invalid date")
}

filled.data$date <- as.Date(filled.data$date)
filled.data$day <- sapply(filled.data$date, FUN = weekday.or.weekend)

averages <- aggregate(steps ~ interval + day, data=filled.data, mean)
ggplot(averages, aes(y= steps, x =interval)) +
        geom_line() +
        facet_grid(day~.) +
        xlab("5-minute interval") + 
        ylab("Number of steps")
        

```